---
title: "Hands-on_Ex07"
format: html
---

# Visualising and Analysing Time-oriented Data

## Getting Started

```{r}
pacman::p_load(scales, viridis, lubridate, ggthemes,
               gridExtra, readxl, knitr, data.table,
               CGPfunctions, ggHoriPlot, tidyverse)
```

### Importing the data

```{r}
attacks <- read_csv("data/eventlog.csv")
```

```{r}
kable(head(attacks))
```

## Plotting Calendar Heatmap

### Data Preparation

Before we can plot the calender heatmap, two new fields namely wkday and hour need to be derived. In this step, we will write a function to perform the task.

```{r}
library(lubridate)
library(data.table)

make_hr_wkday <- function(ts, sc, tz) {

  real_times <- ymd_hms(ts, 
                        tz = tz[1], 
                        quiet = FALSE) 
  
  dt <- data.table(
    source_country = sc,
    # English environment update
    wkday = as.character(lubridate::wday(real_times, label = TRUE, abbr = FALSE, locale = "C")),
    hour = lubridate::hour(real_times)
  )
  return(dt)
}
```

Step 2: Deriving the attacks tibble data frame

```{r}
wkday_levels <- c('Saturday', 'Friday', 
                  'Thursday', 'Wednesday', 
                  'Tuesday', 'Monday', 
                  'Sunday')

attacks <- attacks %>%
  group_by(tz) %>%
  do(make_hr_wkday(.$timestamp, 
                   .$source_country, 
                   .$tz)) %>% 
  ungroup() %>% 
  mutate(wkday = factor(
    wkday, levels = wkday_levels),
    hour  = factor(
      hour, levels = 0:23))
```

Debugging Summary:
The initial issue of a blank ggplot2 tile plot was traced back to an empty data frame caused by the na.omit() function. The root cause lay in the date-time parsing pipeline: because R was running on a system with a Chinese locale, the base function generated weekday names in Chinese (e.g., "星期一"). These failed to match the predefined English factor levels (e.g., "Monday"), causing the entire wkday column to become NA. While attempting to fix this using lubridate to force an English output (locale = "C"), a function masking conflict occurred because data.table also has a wday() function. The ultimate solution was to use explicit namespace calling (lubridate::wday() and lubridate::hour()), which bypassed the conflict, correctly formatted the dates, restored the dataset, and allowed the heatmap to render successfully.

```{r}
kable(head(attacks))
```

### Building the Calendar Heatmaps

```{r}
grouped <- attacks %>% 
  count(wkday, hour) %>% 
  ungroup() %>%
  na.omit()

ggplot(grouped, 
       aes(hour, 
           wkday, 
           fill = n)) + 
geom_tile(color = "white", 
          size = 0.1) + 
theme_tufte(base_family = "sans") + 
scale_fill_gradient(name = "# of attacks",
                    low = "skyblue", 
                    high = "darkblue") +
labs(x = NULL, 
     y = NULL, 
     title = "Attacks by weekday and time of day") +
theme(axis.ticks = element_blank(),
      plot.title = element_text(hjust = 0.5),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6) )
```

###  Building Multiple Calendar Heatmaps

Challenge: Building multiple heatmaps for the top four countries with the highest number of attacks.

Step 1: Deriving attack by country object

In order to identify the top 4 countries with the highest number of attacks, you are required to do the followings:

count the number of attacks by country,
calculate the percent of attackes by country, and
save the results in a tibble data frame.

```{r}
attacks_by_country <- count(
  attacks, source_country) %>%
  mutate(percent = percent(n/sum(n))) %>%
  arrange(desc(n))
```
Step 2: Preparing the tidy data frame

In this step, We are required to extract the attack records of the top 4 countries from attacks data frame and save the data in a new tibble data frame (i.e. top4_attacks).

```{r}
top4 <- attacks_by_country$source_country[1:4]
top4_attacks <- attacks %>%
  filter(source_country %in% top4) %>%
  count(source_country, wkday, hour) %>%
  ungroup() %>%
  mutate(source_country = factor(
    source_country, levels = top4)) %>%
  na.omit()
```

### Plotting Multiple Calendar Heatmaps

Step 3: Plotting the Multiple Calender Heatmap by using ggplot2 package.

```{r}
ggplot(top4_attacks, 
       aes(hour, 
           wkday, 
           fill = n)) + 
  geom_tile(color = "white", 
          size = 0.1) + 
  theme_tufte(base_family = "sans") + 
  coord_equal() +
  scale_fill_gradient(name = "# of attacks",
                    low = "sky blue", 
                    high = "dark blue") +
  facet_wrap(~source_country, ncol = 2) +
  labs(x = NULL, y = NULL, 
     title = "Attacks on top 4 countries by weekday and time of day") +
  theme(axis.ticks = element_blank(),
        axis.text.x = element_text(size = 7),
        plot.title = element_text(hjust = 0.5),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6) )
```

## Plotting Cycle Plot

### Step 1: Data Import
For the purpose of this hands-on exercise, arrivals_by_air.xlsx will be used.

The code chunk below imports arrivals_by_air.xlsx by using read_excel() of readxl package and save it as a tibble data frame called air.

```{r}
air <- read_excel("data/arrivals_by_air.xlsx")
```

### Step 2: Deriving month and year fields
Next, two new fields called month and year are derived from Month-Year field.

```{r}
air$month <- factor(month(air$`Month-Year`), 
                    levels=1:12, 
                    labels=month.abb, 
                    ordered=TRUE) 
air$year <- year(ymd(air$`Month-Year`))
```

### Step 3: Extracting the target country
Next, the code chunk below is use to extract data for the target country (i.e. Vietnam)

```{r}
Vietnam <- air %>% 
  select(`Vietnam`, 
         month, 
         year) %>%
  filter(year >= 2010)
```

### Step 4: Computing year average arrivals by month
The code chunk below uses group_by() and summarise() of dplyr to compute year average arrivals by month. 

```{r}
hline.data <- Vietnam %>% 
  group_by(month) %>%
  summarise(avgvalue = mean(`Vietnam`))
```

### Srep 5: Plotting the cycle plot
The code chunk below is used to plot the cycle plot as shown in Slide 12/23.

```{r}
ggplot() + 
  geom_line(data=Vietnam,
            aes(x=year, 
                y=`Vietnam`, 
                group=month), 
            colour="black") +
  geom_hline(aes(yintercept=avgvalue), 
             data=hline.data, 
             linetype=6, 
             colour="red", 
             size=0.5) + 
  facet_grid(~month) +
  labs(axis.text.x = element_blank(),
       title = "Visitor arrivals from Vietnam by air, Jan 2010-Dec 2019") +
  xlab("") +
  ylab("No. of Visitors") +
  theme_tufte(base_family = "sans")
```

## Plotting Slopegraph

### Step 1: Data Import
Import the rice data set into R environment by using the code chunk below.

```{r}
rice <- read_csv("data/rice.csv")
```

### Step 2: Plotting the slopegraph
Next, code chunk below will be used to plot a basic slopegraph as shown below.

to install CGPfunctions

```{r}
#| message: false
#| warning: false
if (!requireNamespace("remotes", quietly = TRUE)) {
  install.packages("remotes")
}

remotes::install_github("haleyjeppson/ggmosaic", upgrade = "never")

remotes::install_github("ibecav/CGPfunctions", upgrade = "never")

library(CGPfunctions)

```

```{r}
rice %>% 
  mutate(Year = factor(Year)) %>%
  filter(Year %in% c(1961, 1980)) %>%
  newggslopegraph(Year, Yield, Country,
                Title = "Rice Yield of Top 11 Asian Counties",
                SubTitle = "1961-1980",
                Caption = "Prepared by: Dr. Kam Tin Seong")
```