---
title: "Hands-on Exercise 2"
author: "Shuang Zixin"
---

```{r}
#| message: false
#| warning: false

library(pacman)
p_load(tidyverse, ggrepel, ggthemes, hrbrthemes, patchwork)
```

# Beyond ggplot2 Fundamentals
- control the placement of annotation on a graph by using functions provided in ggrepel package,
- create professional publication quality figure by using functions provided in ggthemes and hrbrthemes packages,
- plot composite figure by combining ggplot2 graphs by using patchwork package.

# Getting started

## Installing and loading the required libraries

ggrepel: an R package provides geoms for ggplot2 to repel overlapping text labels.

ggthemes: an R package provides some extra themes, geoms, and scales for ‘ggplot2’.

hrbrthemes: an R package provides typography-centric themes and theme components for ggplot2.

patchwork: an R package for preparing composite figure created using ggplot2.

## Importing data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```
## Beyond ggplot2 Annotation: ggrepel

This code visualizes the correlation between two subjects using geom_point and geom_smooth. Crucially, coord_cartesian is used to enforce the full exam score range (0-100) on the axes, ensuring the context is preserved regardless of the actual data spread.

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label(aes(label = ID), 
             hjust = .5, 
             vjust = -.5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

## Working with ggrepel

This is a refinement of the previous visualization. By swapping geom_label for geom_label_repel (from the ggrepel package), the code automatically adjusts label positions to prevent text from overlapping with data points or other labels, which significantly improves readability in crowded plots.

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

# Beyond ggplot2 Themes

ggplot2 comes with eight built-in themes, they are: theme_gray(), theme_bw(), theme_classic(), theme_dark(), theme_light(), theme_linedraw(), theme_minimal(), and theme_void().

Focusing on visualizing the distribution of a single continuous variable. Explicitly calling theme_gray() ensures the background style is consistent, while the boundary = 100 argument is a nice touch to ensure the bin edges align cleanly with the maximum possible score.

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_gray() +
  ggtitle("Distribution of Maths scores") 
```

## Working with ggtheme package

This illustrates the versatility of the ggthemes package. By simply adding theme_economist(), the plot completely transforms its background, gridlines, and fonts to mimic a professional media publication style, requiring no manual CSS-like adjustments.

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()
```

## Working with hrbthems package

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum()
```

The second goal centers around productivity for a production workflow. In fact, this “production workflow” is the context for where the elements of hrbrthemes should be used. 

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum(axis_title_size = 18,
              base_size = 15,
              grid = "Y")
```

# Beyond Single Graph

It is not unusual that multiple graphs are required to tell a compelling visual story. There are several ggplot2 extensions provide functions to compose figure with multiple graphs. In this section, you will learn how to create composite plot by combining multiple graphs. First, let us create three statistical graphics by using the code chunk below.

```{r}
p1 <- ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") + 
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of Maths scores")
```

next

```{r}
p2 <- ggplot(data=exam_data, 
             aes(x = ENGLISH)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of English scores")
```

Scatterplot for English score versus Maths score

```{r}
p3 <- ggplot(data=exam_data, 
             aes(x= MATHS, 
                 y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

## Creating Composite Graphics: pathwork methods

Patchwork package has a very simple syntax where we can create layouts super easily. Here’s the general syntax that combines:

- Two-Column Layout using the Plus Sign +.
- Parenthesis () to create a subplot group.
- Two-Row Layout using the Division Sign /

## Combining two ggplot2 graphs

```{r}
p1 + p2
```

## Combining three ggplot2 graphs

We can plot more complex composite by using appropriate operators. For example, the composite figure below is plotted by using:

“/” operator to stack two ggplot2 graphs,
“|” operator to place the plots beside each other,
“()” operator the define the sequence of the plotting.


```{r}
(p1 / p2) | p3
```
## Creating a composite figure with tag

In order to identify subplots in text, patchwork also provides auto-tagging capabilities as shown in the figure below.

```{r}
((p1 / p2) | p3) + 
  plot_annotation(tag_levels = 'I')
```

## Creating figure with insert

Beside providing functions to place plots next to each other based on the provided layout. With inset_element() of patchwork, we can place one or several plots or graphic elements freely on top or below another plot.

```{r}
p3 + inset_element(p2, 
                   left = 0.02, 
                   bottom = 0.7, 
                   right = 0.5, 
                   top = 1)
```

## Creating a composite figure by using patchwork and ggtheme

```{r}
patchwork <- (p1 / p2) | p3
patchwork & theme_economist()
```




