---
title: "TakeHome_Ex01"
author: "Shuang Zixin"
---

# **Visualizing and Analysis on Income & Livelihood of Informal Rural Laborers in Vietnam**

# 1. Overview

## 1.1 Research Scene

In the economic structure of developing regions, **informal laborers** play a crucial but often invisible role. Without formal employment contracts or social safety nets, their livelihoods are highly vulnerable to economic fluctuations. This is particularly true in the northern mountainous regions of Vietnam, where limited resources and geographic barriers pose significant challenges.

Visualizing the income sources, asset distribution, and living conditions of these laborers is essential for understanding the economic resilience of rural households and informing targeted policies.

## 1.2 Research Goal

The primary objective of this exercise is to apply visual analytics techniques to reveal patterns, trends, and relationships within the survey data. Specifically, this analysis aims to:

-   **Characterize Income Structures**: Determine the distribution of total earnings and identify income inequality.
-   **Demographic Comparison**: Explore how gender (`CGEN`) and ethnicity (`CRAC`) influence earning potential.
-   **Geographical Analysis**: Visualize the variation in agricultural vs. non-agricultural income across different provinces (`CPRO`).
-   **Financial Health Assessment**: Examine the relationship between household income, debt (`LCRE`), and savings (`LSAV`).
-   **Well-being Correlation**: Investigate whether higher economic status correlates with higher self-reported living satisfaction.

## 1.3 Data Description

The dataset used in this analysis originates from a survey on the livelihoods of informal laborers in rural Vietnam.

::: callout-note
### Data Source

The dataset is retrieved from the study by Hoang & Nguyen (2021).\
**Reference:** [Data from ScienceDirect](https://www.sciencedirect.com/science/article/pii/S235234092100576X#tbl0002)
:::

The raw data contains numerical codes representing various demographic and economic variables. Key variables focused on in this study include:

-   **Demographics**: `CPRO` (Province), `CGEN` (Gender), `CRAC` (Race/Ethnicity).
-   **Financials**: `TEIN` (Total Earned Income), `TAIN` (Total Agricultural Income), `TOIN` (Total Non-farm Income).
-   **Assets & Debt**: `LCRE` (Loan/Credit), `LSAV` (Savings).
-   **Perception**: `LHO` (Living Condition Assessment).

# 2. Data Preparation

In this section, we will load the necessary R packages, import the dataset, and perform data wrangling to transform raw numerical codes into meaningful categorical labels suitable for analysis.

## 2.1 Loading Packages

First, we load the required packages. We will use `pacman` for efficient package management. The core package is `tidyverse` for data manipulation and visualization, supplemented by `ggrepel` for label formatting, `ggthemes` for plot aesthetics, and `knitr` for generating tables.

```{r}
#| message: false
#| warning: false
#| class-output: output-r

library(pacman)
p_load(tidyverse,    # Core data science packages (ggplot2, dplyr, readr, etc.)
       ggrepel,      # Repulsive text labels
       ggthemes,     # Extra themes for ggplot2
       hrbrthemes,   # Typography-centric themes
       patchwork,    # For combining plots
       knitr,        # For static tables
       kableExtra,   # For styling tables
       readxl)       # For imporitng excel file
```

## 2.2 Data Importing

The dataset is imported using read_excel. We display the first few rows to inspect the raw structure.

```{r}
#| message: false
#| warning: false
#| class-output: output-r

raw_data <- read_excel("data/Data.xlsx")
```

The newly imported dataset consists of **725 rows** and **30 columns**. The detailed data structure can be viewed in the collapsible section below.

::: {.callout-note collapse="true"}
```{r}
glimpse(raw_data)
```
:::

## 2.3 Data Wrangling

The raw dataset contains numerical codes for categorical variables and requires significant preprocessing to be suitable for visualization. To ensure data accuracy and interpretability, we implemented a five-step data wrangling pipeline:

### Step 1: Variable Selection and Renaming

The original dataset contains 30 columns. We narrowed our scope to key variables relevant to the analysis objectives: \* **Demographics**: Province, Gender, Ethnicity, Job Type. \* **Financials**: Income components (Total, Agricultural, Service/Industrial, Others), Debt (`LCRE`), and Savings (`LSAV`). \* **Living Standards**: House status indicators (`LHO1` - `LHO4`). \* **Factors**: Education Level (`FEDU`).

All selected columns were renamed from cryptic codes (e.g., `CGEN`, `TEIN`) to descriptive English labels (e.g., `Gender`, `Income_Total`).

```{r}
#| class-output: output-r
#| message: false
#| warning: false

clean_data <- raw_data %>%
  select(
    CPRO, CGEN, CRAC, CJOB,       # Demographics
    TEIN, TAIN, TSII, TOIN,       # Income (Added TSII)
    FEDU,                         # Education
    LCRE, LSAV,                   # Financials
    LHO1, LHO2, LHO3, LHO4        # Housing Conditions
  ) %>%
  rename(
    Province        = CPRO,
    Gender          = CGEN,
    Ethnicity       = CRAC,
    Job_Type        = CJOB,
    Income_Total    = TEIN,
    Income_Agri     = TAIN,
    Income_Service  = TSII,      # New Column
    Income_Other    = TOIN,
    Education_Level = FEDU,
    Debt            = LCRE,
    Savings         = LSAV
  )
```

### Step 2: Handling Missing Values (Imputation)

We applied specific imputation strategies based on the variable definitions in Table 1:

-   **Quantitative Variables (Income, Debt, Savings):** According to Table 1, variables like `TEIN` (Total Income) and `LCRE` (Loan) are defined as **"Scale"** measures. In this survey context, a missing value (`NA`) in a component income field (e.g., Service Income) logically implies a zero contribution to the total income, rather than missing data. Therefore, we imputed these `NA` values as `0` to maintain mathematical consistency in total income calculations.

-   **Categorical Variables (Job Type):** Conversely, for the `Job_Type` variable, Table 2 explicitly lists a "Missing System" category ($n=49$). We preserved these as a distinct "Missing/Unknown" category instead of imputing them.

```{r}
clean_data <- clean_data %>%
  mutate(
    across(c(Debt, Savings, Income_Agri, Income_Service, Income_Other), ~replace_na(., 0))
  ) 
```

### Step 3: Recoding Categorical Variables (Verification & Mapping)

Since the dataset lacks an embedded codebook, we reconstructed the coding schema by cross-referencing the internal frequency counts with the summary statistics provided in **Table 2** and **Table 3** of the source article (Hoang & Nguyen, 2021). This "frequency matching" process confirmed the following mappings:

-   **Job Type (`CJOB`)**: Code `1` appears 515 times, which perfectly matches the count for "Agricultural sector" in Table 2. Thus, `1` = Agriculture, `2` = Service, `3` = Others.
-   **Education (`FEDU`)**: Code `0` appears 473 times, matching the "Primary education" count in Table 3. Code `1` (n=136) matches "Lower secondary". This corrects the initial assumption that 0 represented "No Schooling".
-   **House Type (`LHO`)**: The columns `LHO1` to `LHO4` were identified as dummy variables representing housing quality, ranging from "Cottage" (LHO1) to "Buildings" (LHO4).

```{r}
clean_data <- clean_data %>%
mutate(
    Province = case_match(Province,
      1 ~ "Province 1", 2 ~ "Province 2", 3 ~ "Province 3",
      4 ~ "Province 4", 5 ~ "Province 5",
      .default = as.character(Province)
    ),
    
    Gender = case_match(as.character(Gender),
      "1" ~ "Male",
      "2" ~ "Female",
      .default = NA_character_
    ),
    
    Ethnicity = case_match(Ethnicity,
      1 ~ "Kinh",
      2 ~ "Ethnic Minorities"
    ),
    
    Job_Type = case_match(Job_Type,
      1 ~ "Agriculture",
      2 ~ "Service & Ind", 
      3 ~ "Others",
      NA ~ "Missing/Unknown",
      .default = "Missing/Unknown"
    ),
    
    Education_Level = case_match(Education_Level,
      0 ~ "Primary",
      1 ~ "Lower Secondary",
      2 ~ "Upper Secondary",
      3 ~ "Others",
      .default = "Primary" # Default to largest group if unknown
    )
  )
```

### Step 4: Feature Engineering

To deepen our analysis, we derived new variables from the raw data:

**`House_Type`**: Consolidated the four `LHO` dummy columns into a single categorical variable to represent the highest quality of housing owned.

**`Financial_Status`**: Categorized households into "In Debt" or "Debt Free" based on the `Debt` column.

**`Livelihood_Group`**: Classified households based on their income structure (e.g., "Farming Focused" if agricultural income constitutes \>50% of total income).

```{r}
clean_data <- clean_data %>%
  mutate(
    # A. Housing Quality (Consolidating LHO1-4 into one column)
    # Logic: Check which column is 1. Higher number = Better house.
    House_Type = case_when(
      LHO4 == 1 ~ "Buildings (Best)",
      LHO3 == 1 ~ "Solid House",
      LHO2 == 1 ~ "Roofed House",
      LHO1 == 1 ~ "Cottage (Basic)",
      TRUE      ~ "Unknown"
    ),
    
    # B. Financial Health
    Financial_Status = if_else(Debt > 0, "In Debt", "Debt Free"),
    
    # C. Livelihood Strategy (Income Source)
    # Calculate percentages
    Pct_Agri = (Income_Agri / Income_Total) * 100,
    Pct_Service = (Income_Service / Income_Total) * 100,
    
    Livelihood_Group = case_when(
      Pct_Agri >= 50 ~ "Farming Focused",
      Pct_Service >= 50 ~ "Service/Ind Focused",
      TRUE ~ "Diversified/Other"
    )
  )
```

### Step 5: Formatting

Finally, all categorical variables were converted into **factors**. Ordinal variables such as `Education_Level` and `House_Type` were explicitly ordered to respect their logical hierarchy (e.g., Primary \< Secondary) for correct plotting.

```{r}
clean_data <- clean_data %>%
  
  select(-c(LHO1, LHO2, LHO3, LHO4)) %>%
  
  mutate(
    across(c(Province, Gender, Ethnicity, Financial_Status, Livelihood_Group), as.factor),
    
    Job_Type = factor(Job_Type, 
                      levels = c("Agriculture", "Service & Ind", "Others", "Missing/Unknown")),
    
    Education_Level = factor(Education_Level, 
                             levels = c("Primary", "Lower Secondary", "Upper Secondary", "Others")),
    
    House_Type = factor(House_Type, 
                        levels = c("Cottage (Basic)", "Roofed House", "Solid House", "Buildings (Best)"))
  )
```

### Data Inspection

After completing the cleaning and feature engineering pipeline, the final dataset consists of 725 observations and 16 variables (excluding intermediate columns). The table below provides a preview of the processed data structure, confirming that all variables are correctly formatted and ready for visual analysis.

::: {.callout-note collapse="true"}
```{r}
glimpse(clean_data)
```
:::

# 3. Visualization & Analysis

## 3.1 Income Distribution

### 3.1.1 Overview & Methodology

In this section, we examine the distribution of Total Household Income (`Income_Total`) to understand the economic landscape of the surveyed population.

-   **Variable Type**: Quantitative Continuous (`Income_Total`) and -Nominal (`Province`).

-   **Visualisation Technique**: A Histogram with Density Overlay, faceted by Province.

-   **Rationale**: Income data is typically non-normal. The histogram allows us to visualize the frequency of income ranges (bins), while the overlaid density curve provides a smoothed approximation of the distribution shape. Faceting by `Province` enables a direct comparison of economic disparity across different geographic regions.

### 3.1.2 Visual Analysis

::: panel-tabset
## Code

```{r}
#| label: fig-income-dist
#| fig-cap: "Distribution of Total Household Income by Province"
#| message: false
#| warning: false
#| eval: false
#| echo: true

median_income <- clean_data %>%
  group_by(Province) %>%
  summarise(grp_median = median(Income_Total, na.rm = TRUE))

ggplot(clean_data, aes(x = Income_Total)) +
  geom_histogram(aes(y = after_stat(density)), 
                 binwidth = 10, 
                 fill = "#69b3a2", 
                 color = "white", 
                 alpha = 0.7) +

  geom_density(color = "#e63946", lwd = 1) +

  geom_vline(data = median_income, aes(xintercept = grp_median), 
             color = "darkred", linetype = "dashed", size = 0.8) +

  facet_wrap(~Province, ncol = 3) +

  theme_minimal() +
  labs(
    title = "Distribution of Household Income across Provinces",
    subtitle = "Red dashed line represents the Median Income",
    x = "Total Income (Million VND)",
    y = "Density"
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    strip.text = element_text(face = "bold") 
  )
```

## Graph
```{r}
#| label: fig-income-render
#| ref.label: "fig-income-dist"
#| echo: false
#| eval: true
#| message: false
#| warning: false
```
:::

### 3.1.3 Observations

The visualization reveals a striking similarity in income distribution across all five provinces, suggesting a regionally homogenous economic structure rather than localized disparities.

- **Distribution Shape**: All five provinces exhibit identical right-skewed distributions, indicating a uniform concentration of households at lower income levels across the region.
- **Central Tendency**: Median incomes show minimal variation (range: 20-35 million VND), suggesting that geographic location is not a primary driver of household wealth in this dataset.

## 3.2 Demographic Disparities

### 3.2.1 Overview & Methodology

This section investigates how social stratification—specifically **Ethnicity** and **Gender**—correlates with economic outcomes.

-   **Variable Type**: Quantitative Continuous (`Income_Total`) vs. Categorical Nominal (`Ethnicity`, `Gender`).
-   **Visualisation Technique**: A **Grouped Boxplot**.
-   **Rationale**: Boxplots are the statistical standard for comparing distributions across groups. They clearly display the **Median** (central tendency), **Interquartile Range (IQR)** (wealth dispersion), and **Outliers** (extreme values). Grouping by Ethnicity and filling by Gender allows us to detect intersectional disparities (e.g., do female-headed households fare worse within ethnic minority groups?).

### 3.2.2 Visual Analysis

::: panel-tabset
## Code

```{r}
#| label: fig-demo-disparity
#| fig-cap: "Income Disparities by Ethnicity and Gender of Household Head"
#| message: false
#| warning: false
#| eval: false
#| echo: true

ggplot(clean_data, aes(x = Ethnicity, y = Income_Total, fill = Gender)) +

  geom_boxplot(alpha = 0.8, outlier.colour = "red", outlier.shape = 1) +
  
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, color = "white", 
               position = position_dodge(0.75)) +

  scale_y_continuous(trans = "log10", labels = scales::comma, breaks = c(1, 10, 50, 100, 500)) +
  
  scale_fill_manual(values = c("Male" = "#4e79a7", "Female" = "#f28e2b")) +
  theme_minimal() +
  
  labs(
    title = "Income Inequality: Ethnicity vs. Gender",
    subtitle = "White diamonds indicate Mean Income; Red circles indicate Outliers",
    x = "Ethnic Group",
    y = "Total Income (Million VND)",
    fill = "Head of Household"
  ) +
  theme(
    legend.position = "top",
    axis.text.x = element_text(face = "bold", size = 11),
    plot.title = element_text(face = "bold", size = 14)
  )

```

## Graph
```{r}
#| label: fig-demo-render
#| ref.label: "fig-demo-disparity"
#| echo: false
#| eval: true
#| message: false
#| warning: false
```

## Statistics

```{r}
#| echo: false
#| message: false

library(knitr)

stats_table <- clean_data %>%
  group_by(Ethnicity) %>%
  summarise(
    Median = median(Income_Total, na.rm = TRUE),
    Mean = mean(Income_Total, na.rm = TRUE),
    Count = n()
  ) %>%
  mutate(
    `Gap vs Minority (%)` = round((Median - 28.5) / 28.5 * 100, 1) 
  )

kable(stats_table, caption = "Table: Summary Statistics of Income by Ethnicity", digits = 1)
```
:::

### 3.2.3 Observations

The analysis reveals a nuanced picture of inequality where widespread poverty coexists with a distinct structural gap.

- **Structural Disparity**: The Kinh group median income (35.0m VND) is 22.8% higher than that of Ethnic Minorities (28.5m VND), confirming a distinct economic advantage.
- **Upper-End Divergence**: While low-income prevalence is similar between groups, the wider gap in Mean Income (41.6m vs 34.9m) indicates that high-income outliers are predominantly Kinh.


## 3.3 Geographical Nuances

### 3.3.1 Overview & Methodology

This section examines the **economic efficiency** of different livelihood strategies across provinces.

-   **Variable Interaction**: `Province` (Nominal) × `Livelihood_Group` (Nominal) → `Median Income` (Ratio).
-   **Visualisation Technique**: **Heatmap**.
-   **Rationale**: A heatmap provides an immediate visual assessment of "what pays best where." By mapping median income to color intensity within a matrix grid, we can instantly identify high-efficiency economic hotspots versus low-yield traps without the clutter of complex distribution plots.

### 3.3.2 Visual Analysis

::: panel-tabset
## Code

```{r}
#| label: fig-geo-heatmap
#| fig-cap: "Economic Efficiency Heatmap: Median Income by Strategy and Region"
#| message: false
#| warning: false
#| eval: false
#| echo: true

library(dplyr)
library(ggplot2)

heatmap_data <- clean_data %>%
  group_by(Province, Livelihood_Group) %>%
  summarise(
    Median_Income = median(Income_Total, na.rm = TRUE),
    .groups = 'drop'
  )

ggplot(heatmap_data, aes(x = Province, y = Livelihood_Group, fill = Median_Income)) +
  
  geom_tile(color = "white", lwd = 1.5, linetype = 1) +
  

  geom_text(aes(label = round(Median_Income, 1)), color = "white", size = 4, fontface = "bold") +
  

  scale_fill_viridis_c(option = "magma", direction = -1, name = "Median Income\n(Million VND)") +
  
  scale_x_discrete(expand = c(0, 0)) + 
  scale_y_discrete(expand = c(0, 0)) + 
  theme_minimal() +
  labs(
    title = "Regional Economic Efficiency Matrix",
    subtitle = "Heatmap showing Median Income",
    x = NULL,
    y = NULL
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 10),
    axis.text = element_text(size = 6, face = "bold", color = "gray30"),
    legend.position = "right",
    panel.grid = element_blank() 
  )

```

## Graph
```{r}
#| label: fig-geo-render
#| ref.label: "fig-geo-heatmap"
#| echo: false
#| eval: true
#| message: false
#| warning: false
```
:::

### 3.3.3 Observations

The heatmap data indicates three key patterns regarding income generation:

- **Sector Performance**: Households relying solely on farming consistently record the lowest median incomes across all provinces.
- **Diversification Return**: 'Wage' and 'Diversified' categories correlate with significantly higher incomes than farming, with Province 1 exhibiting the strongest non-farm returns.

## 3.4 Education Level and Household Income

### 3.4.1 Overview & Methodology

This section analyzes the relationship between the education level of the household head and total household income.

-   **Variables**: `Income_Total` (Ratio) grouped by `Education_Level` (Ordinal/Nominal).
-   **Visualisation**: **Empirical Cumulative Distribution Function (ECDF)**.
-   **Rationale**: The ECDF plots the cumulative percentage of households earning up to a certain income level. This allows for a comparison of the entire income distribution across education groups. A curve positioned further to the right indicates higher income levels across the population segment.

### 3.4.2 Visual Analysis

::: panel-tabset
## Code

```{r}
#| label: fig-edu-ecdf
#| fig-cap: "Cumulative Income Distribution by Education Level"
#| message: false
#| warning: false
#| eval: false
#| echo: true

library(ggplot2)
library(dplyr)
library(scales)

cdf_data <- clean_data %>%

  filter(!is.na(Education_Level) & !is.na(Income_Total)) %>%

  mutate(Education_Level = factor(Education_Level, 
                                  levels = c("Primary", "Lower Secondary", "Upper Secondary", "Others")))

ggplot(cdf_data, aes(x = Income_Total, color = Education_Level)) +
  
  stat_ecdf(geom = "step", linewidth = 1) +
  
  scale_y_continuous(labels = scales::percent, name = "Cumulative Proportion of Households") +
  
  scale_color_brewer(palette = "Set1", name = "Education Level") +
  
  geom_vline(xintercept = 20, linetype = "dashed", color = "gray60") +
  
  theme_minimal() +
  labs(
    title = "Income Distribution by Education Level (ECDF)",
    x = "Total Household Income (Million VND)"
  ) +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold", size = 10)
  )
```

## Graph
```{r}
#| label: fig-edu-render
#| ref.label: "fig-edu-ecdf"
#| echo: false
#| eval: true
#| message: false
#| warning: false
```
:::

### 3.4.3 Observations

The ECDF analysis indicates a strong correlation between education levels and household income stability:

- **Poverty Reduction**: At the 20 million VND threshold, 30% of Primary-educated households fall below the line, compared to under 5% of the Upper Secondary group.
- **Marginal Return**: The income gap between Lower and Upper Secondary levels exceeds that of earlier stages, indicating the highest marginal gain occurs upon completing Upper Secondary education.

## 3.5 Living Standards: Housing Conditions

### 3.5.1 Overview & Methodology

This section analyzes the income distribution across different housing conditions.

-   **Variables**: `Income_Total` (Ratio) grouped by `House_Type` (Ordinal).
-   **Visualisation**: **Ridgeline Plot (Density Plot)**.
-   **Rationale**: The ridgeline plot visualizes the entire probability density of income for each housing category. Unlike boxplots which summarise data into quartiles, this method reveals the shape of the distribution—showing peaks (most common income levels) and tails (wealthy outliers)—providing a richer understanding of the economic reality behind housing choices.

### 3.5.2 Visual Analysis

::: panel-tabset
## Code

```{r}
#| label: fig-housing-ridge
#| fig-cap: "Income Distribution by Housing Type (Ridgeline Plot)"
#| message: false
#| warning: false
#| eval: false
#| echo: true

library(ggplot2)
library(ggridges)
library(dplyr)
library(scales)

ridge_data <- clean_data %>%

  filter(!is.na(Income_Total) & !is.na(House_Type)) %>%
  filter(House_Type != "Unknown") %>%

  add_count(House_Type) %>%
  filter(n > 5) %>%
  
  mutate(House_Type = factor(House_Type, 
                             levels = c("Cottage (Basic)", "Roofed House", "Solid House", "Buildings (Best)")))


ggplot(ridge_data, aes(x = Income_Total, y = House_Type, fill = House_Type)) +
  
  geom_density_ridges(
    scale = 1.2,          
    rel_min_height = 0.01, 
    alpha = 0.7,           
    quantile_lines = TRUE, 
    quantiles = 2         
  ) +

  scale_x_log10(labels = scales::comma, breaks = c(10, 20, 50, 100, 200, 500)) +

  scale_fill_brewer(palette = "Spectral", direction = -1) +

  theme_minimal() +
  labs(
    title = "Income Densities by Housing Type",
    subtitle = "Vertical lines indicate median income for each group",
    x = "Total Household Income (Million VND) [Log Scale]",
    y = NULL,
    fill = "Housing Type"
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    legend.position = "none"
  )
```

## Graph
```{r}
#| label: fig-housing-render
#| ref.label: "fig-housing-ridge"
#| echo: false
#| eval: true
#| message: false
#| warning: false
```
:::

### 3.5.3 Observations

The Ridgeline plot indicates the following specific relationships between housing type and income distribution:

- **Income Gradient**: Density peaks shift rightward from "Cottage" (low income) to "Buildings" (high income), validating housing quality as a wealth proxy.
- **Variance by Type**: "Cottage" households exhibit low variance (narrow peak at 15-20m VND), whereas "Buildings" show high variance (flat distribution), indicating greater income heterogeneity among the wealthy.



# 4. Conclusion

### 4.1 Drivers of Income Inequality

The comparative analysis of five key variables identifies demographic and structural factors, rather than geographic location, as the primary determinants of household income.

- **Demographic over Geographic**: While income distributions remain uniform across all five provinces (median range: 20-35m VND), significant stratification exists between ethnic groups. The Kinh majority maintains a 22.8% higher median income compared to Ethnic Minorities.
- **The Education Threshold**: Human capital accumulation shows non-linear returns. The marginal income increase is minimal between Primary and Lower Secondary levels but expands significantly upon completion of Upper Secondary education. This identifies Upper Secondary certification as a critical differentiator for entering higher income brackets.

### 4.2 Livelihood and Asset Accumulation

Income disparities are further reinforced by livelihood strategies and material living standards.

- **Sectoral Ceilings**: Households relying exclusively on agriculture consistently record the lowest median incomes. Higher income levels are associated with 'Wage' and 'Diversified' livelihood strategies, confirming that transition out of pure farming is a prerequisite for significant income growth.
- **Asset Validation**: Housing quality serves as a reliable proxy for economic status. The low variance in income among "Cottage" dwellers (concentrated at 15-20m VND) contrasts with the high variance among "Building" residents. This confirms that while poverty is structurally uniform (low income, poor housing), wealth accumulation is more heterogeneous.

### 4.3 Summary

The data characterizes the region as geographically homogenous but structurally divided. Low income is systematically linked to three converging factors: **Ethnic Minority status, low education (Lower Secondary or below), and reliance on pure agriculture.** Conversely, higher economic performance is driven by the completion of Upper Secondary education and engagement in the non-farm economy.

# 5. references

### Data Manipulation & Cleaning (`dplyr`)

https://r4ds.hadley.nz/data-transform
- *Used for:* `filter()`, `mutate()`, and `group_by()` functions.

### Data Visualization (`ggplot2`)

https://ggplot2.tidyverse.org
- *Used for:* The foundational "Grammar of Graphics" structure used in all plots.


